# my little reminder for (just some) time aggregation methods
### obviously many others are also possible but not here contemplated (known)!

*****

## generate some fake data to play with

#### time sequence using POSIXct class (storing date-time and many other features like tz, dst); the genreal rule of thumb is to stick with the simplest date/time class one needs (...and that's why I'm keeping up with the most complex and dangeorus one, ugh!) 

```{r}
# start and end of time seq
start <- as.POSIXct('2014-05-01 00:00:00')
end <- as.POSIXct('2014-05-02 00:00:00')

# by sec
bysec<-as.POSIXct(seq(start, end, by="sec"))
#alternatively   
bysec<-as.POSIXct(seq(start, end, by=1))

# by min
bymin<-as.POSIXct(seq(start, end, by="min"))
# alternatively   
bymin<-as.POSIXct(seq(start, end, by=60))

# by hour
byhour<-as.POSIXct(seq(start, end, by="hour"))
# alternatively 
byhour<-as.POSIXct(seq(start, end, by=60*60))

# and so on for other time intervals...

# eg. by 15 minutes
by15min<-as.POSIXct(seq(start, end, by="15 mins"))
# alternatively   
by15min<-as.POSIXct(seq(start, end, by=60*15))

```

#### build up a df with 3 vars using "bymin" time sequence

```{r}
set.seed(123)
v1<-rnorm(length(bymin), 5)

set.seed(456)
v2<-rnorm(length(bymin), 10)

# here using bymin as time seq
df<-data.frame(date=bymin,v1, v2)

# introduce some random NAs
df[sample(nrow(df),20),"v1"]<-NA
df[sample(nrow(df),10),"v2"]<-NA

summary(df)
str(df)
```

## aggregation through base methods (no packages)
****

### mean by hour using format
```{r}
head(aggregate(df[c("v1","v2")], format(df["date"],"%Y-%m-%d %H"), mean, na.rm = TRUE))
#alternatively
head(aggregate(df[c(2,3)], format(df[1],"%Y-%m-%d %H"), mean, na.rm = TRUE))
```

### mean by hour using cut
```{r}
# see cut.POSIXt for refs
head(aggregate(df[c("v1","v2")], list(date=cut(df$date, breaks="hour")), mean, na.rm = TRUE))
```

### mean by 15 mins using cut
```{r}
head(aggregate(df[c("v1","v2")], list(date=cut(df$date, breaks="15 mins")), mean, na.rm = TRUE))
```

## aggregation through some dedicated package methods
****

## library zoo (just some relvant examples)

```{r}
require(zoo)
```
### myzoo object
```{r}
myzoo<-zoo(cbind(df$v1,df$v2),df$date)
str(myzoo)
```
#### note index() or time()

### plot zoo series with lattice
```{r}
require(lattice)
xyplot(myzoo, screens=c("v1","v2"),
       scales = list(y = list(relation = "same"))
       )
xyplot(myzoo, superpose=TRUE, auto.key=list(text=c("v1", "v2")))

```

### mean by hour using format

```{r}
head(aggregate(myzoo, format(index(myzoo),"%Y-%m-%d %H"), mean, na.rm=TRUE))
```


### mean by 15 mins using cut

```{r}
head(aggregate(myzoo, cut(index(myzoo), breaks="15 mins"), mean, na.rm=TRUE))
```

### rolling average by 5 mins (trailing average)

```{r}
head(rollapply(myzoo,width=5,mean, align="right", na.rm=TRUE))
```
#### note that rollmean() does not handle NAs

### rolling average by 5 mins (average at every 5 mins)

```{r}
head(rollapply(myzoo,width=5,FUN=mean, by=5,align="right", na.rm=TRUE))
```


## library xts
#### pretty similar to zoo object (derived by) but with some dedicated useful functions
```{r}
require(xts)
```
### myxts object
```{r}
myxts<-xts(cbind(df$v1,df$v2),df$date)
str(myxts)
```
### mean by 15 mins
#### to note all dates are aligned to the end of each period by default
```{r}
head(period.apply(myxts,endpoints(myxts, "mins", 15), mean, na.rm=TRUE))
```

### mean by 2 hours mins
```{r}
head(period.apply(myxts,endpoints(myxts, "hours", 2), mean, na.rm=TRUE))
```
### xts has some useful facilities for subsetting
```{r}
#periodicity of dataset
periodicity(myxts)

#subset one single record
myxts["2014-05-01 08:00"]

#subset starting from first rec up to...
myxts["/2014-05-01 00:05"]

#subset first 5 recs
first(myxts, "5 mins")

#subset last 5 recs
last(myxts, "5 mins")

```
