# my little reminder...
***
# ...for some time aggregation methods in R
***

### obviously many other methods are possible but not here contemplated (...and also not known!)

*****

## generate some fake data to play with


time sequence using POSIXct class (storing date-time and many other features like tz, dst);   
general rule of thumb is to stick with the simplest date/time class one needs (...and that's why I'm keeping up with the most complex and dangeorus one, ugh!) 

```{r}
# start and end of time seq
start <- as.POSIXct('2014-05-01 00:00:00')
end <- as.POSIXct('2014-05-02 00:00:00')

# by sec
bysec<-as.POSIXct(seq(start, end, by="sec"))
#alternatively   
bysec<-as.POSIXct(seq(start, end, by=1))

# by min
bymin<-as.POSIXct(seq(start, end, by="min"))
# alternatively   
bymin<-as.POSIXct(seq(start, end, by=60))

# by hour
byhour<-as.POSIXct(seq(start, end, by="hour"))
# alternatively 
byhour<-as.POSIXct(seq(start, end, by=60*60))

# and so on for other time intervals...

# eg. by 15 minutes
by15min<-as.POSIXct(seq(start, end, by="15 mins"))
# alternatively   
by15min<-as.POSIXct(seq(start, end, by=60*15))
```

build up a df with 3 vars using "bymin" time sequence

```{r}
set.seed(123)
v1<-rnorm(length(bymin), 5)

set.seed(456)
v2<-rnorm(length(bymin), 10)

# here using bymin as time seq
df<-data.frame(date=bymin,v1, v2)

# introduce some random NAs
df[sample(nrow(df),20),"v1"]<-NA
df[sample(nrow(df),10),"v2"]<-NA

summary(df)
str(df)
```
****
## time aggregation through base methods (no packages)
****

### mean by hour using format
```{r}
head(aggregate(df[c("v1","v2")], format(df["date"],"%Y-%m-%d %H"), mean, na.rm = TRUE))
#alternatively
head(aggregate(df[c(2,3)], format(df[1],"%Y-%m-%d %H"), mean, na.rm = TRUE))
```

### mean by hour using cut
```{r}
# see cut.POSIXt for refs
head(aggregate(df[c("v1","v2")], list(date=cut(df$date, breaks="hour")), mean, na.rm = TRUE))
```

### mean by 15 mins using cut
```{r}
head(aggregate(df[c("v1","v2")], list(date=cut(df$date, breaks="15 mins")), mean, na.rm = TRUE))
```

### plot means by hour
```{r}

#aggregate data by hour
means<-aggregate(df[c("v1","v2")], format(df["date"],"%Y-%m-%d %H"), mean, na.rm = TRUE)
str(means)

#create x axis temporal sequence (POSIXct) to be plotted
means$date<-seq(df$date[1],df$date[nrow(df)], length=nrow(means))
str(means)

```

NB: many other formatting (aggregating) time sequences are also possible, see ?strptime for reference to further abbreviations

```{r}
# default plot
plot(means$date, means$v1, type="l")
```

```{r}
# default axis (number of tickmarks) with user formatted labels
plot(means$date, means$v1, type="l", xaxt="n")
axis.POSIXct(1, means$date, format="hh %H")
```

```{r}
# user defined axis (number of tickmarks) and formatted labels 
plot(means$date, means$v1, type="l", xaxt="n")
axis.POSIXct(1, means$date, at=seq(min(means$date), max(means$date), by="hours"),
             format="hh %H", las=2, cex=0.8)
# adding some reference lines
abline(v=as.POSIXct("01/05/2014 00", format="%d/%m/%Y %H"), col="cyan")
abline(v=as.POSIXct("01/05/2014 12", format="%d/%m/%Y %H"), col="red")
abline(v=as.POSIXct("02/05/2014 00", format="%d/%m/%Y %H"), col="green")
```

****
## time aggregation through some dedicated package methods
****

## library zoo (just some relevant examples)

```{r}
require(zoo)
```

### myzoo object

```{r}
myzoo<-zoo(cbind(df$v1,df$v2),df$date)
str(myzoo)
```

NB: index() or time()

### plot zoo series with lattice

```{r}
require(lattice)
# default plot
xyplot(myzoo, screens=c("v1","v2"),
       scales = list(y = list(relation = "same"))
       )
# superposed plot
xyplot(myzoo, superpose=TRUE, auto.key=list(text=c("v1", "v2")))

# user defined axis (number of tickmarks) and formatted labels 
mydate<-seq(from = min(index(myzoo)), to = max(index(myzoo)), length.out = 6)

xyplot(myzoo, superpose=TRUE, auto.key=list(text=c("v1", "v2")),
       scale = list(x = list(at = mydate, labels = format(mydate, "hh %H"),
                             cex = 0.8, rot=90))
       )
#alternative
mydate2<-seq(from = min(index(myzoo)), to = max(index(myzoo)), by="4 hours")
xyplot(myzoo, superpose=TRUE, auto.key=list(text=c("v1", "v2")),
       scale = list(x = list(at = mydate2, labels = format(mydate2, "hh %H"),
                             cex = 0.8, rot=90))
       )
```

### mean by hour using format

```{r}
head(aggregate(myzoo, format(index(myzoo),"%Y-%m-%d %H"), mean, na.rm=TRUE))
```

### mean by 15 mins using cut

```{r}
head(aggregate(myzoo, cut(index(myzoo), breaks="15 mins"), mean, na.rm=TRUE))
```

### rolling average by 5 mins (trailing average)

```{r}
head(rollapply(myzoo,width=5,mean, align="right", na.rm=TRUE))
```

NB: rollmean() does not handle NAs

### rolling average by 5 mins (average at every 5 mins)

```{r}
head(rollapply(myzoo,width=5,FUN=mean, by=5,align="right", na.rm=TRUE))
```


## library xts

pretty similar to zoo object (derived by) but with some dedicated useful functions

```{r}
require(xts)
```

### myxts object

```{r}
myxts<-xts(cbind(df$v1,df$v2),df$date)
str(myxts)
```

### mean by 15 mins

NB: all dates are aligned to the end of each period by default

```{r}
head(period.apply(myxts,endpoints(myxts, "mins", 15), mean, na.rm=TRUE))
```

### mean by 2 hours mins

```{r}
head(period.apply(myxts,endpoints(myxts, "hours", 2), mean, na.rm=TRUE))
```
### xts has some useful facilities for subsetting

```{r}
#periodicity of dataset
periodicity(myxts)

#subset one single record
myxts["2014-05-01 08:00"]

#subset starting from first rec up to...
myxts["/2014-05-01 00:05"]

#subset first 5 recs
first(myxts, "5 mins")

#subset last 5 recs
last(myxts, "5 mins")
```
